### 2019-07-04 2-Chapters/2-Chapter/chapter2.tex
Using uniform notations for Exp3, Exp3.S, Exp3.R

Similarly to what is considered for the first lower-bound,
a natural question is to know if there is an algorithm achieving a regret upper-bound of the form $R_T^{\cA}(I) \leq \bigO{\sqrt{K T}}$ for any instance, independently on the problem difficulty.
The question has been answered since the early $2000$s, as it is for instance the case for the Exp3 algorithm from \cite{Auer02}.
In the last decade, some algorithms were shown to achieve both a problem-dependent logarithmic and a minimax upper-bounds,
like MOSS from \cite{Audibert2009minimax} or recently \KLUCBpp{} from \cite{Menard17},
and such results are usually referred to as ``best of both worlds''.

@@

% ----------------------------------------------------------------------------
\textbf{Policies for adversarial bandits.}
%
Another important family of bandits algorithms are the algorithms proposed for the adversarial setting \cite{Auer02NonStochastic}.
The first of them is the Exp3 policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loration'',
which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details, see
% on Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
this article by J. Kun, \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-}}\\
\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{of-the-multiplicative-weights-update-algorithm}}. The survey \cite{Arora2012multiplicative} gives more details.},
% which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019)
a family of algorithms developed since the 1950s, and applied to a wide range of different domains.
% https://smpybandits.github.io/docs/Policies.Exp3.html
%
The Exp3 policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution, \ie,
$\Pr(A(t)=k) \eqdef (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
The weights are initially uniform, \ie, $w_k(0)=1/K$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, Exp3 updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$. The weights are then renormalized, to have a sum $\sum_k w_k(t+1) = 1$.
This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r(t)=Y_{k,t}$, that is, $\widehat{r_k}(t) = r(t) / \Pr(A(t)=k)$.
%
Like for the $\varepsilon$-greedy policy, the parameter $\gamma$ can be constant, or can be a non-increasing sequence $(\gamma_t)_{t\in\N^*}$.
%
It was proven in \cite{Auer02NonStochastic} that Exp3 with constant parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
while using a decreasing sequence, such as $\gamma_t = \sqrt{\log(K) / K} \log(t) / t$ gives an order-optimal problem-independent regret upper-bound of $\bigO{\sqrt{K T}}$.
A short proof of this results and more details on Exp3 are given in \cite{Bubeck12}.
%
Many other policies have been proposed for adversarial bandits, like Follow-the-Perturbed-Leader or Follow-the-Regularized-Leader for linear adversarial bandits,
% but we do not study them in this thesis,
and we refer to Parts~III and VI of \cite{LattimoreBanditAlgorithmsBook} for more details.

@@

% https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
Finally, it is worth mentioning the Exp3$^{++}$ algorithm, as a recent variant of Exp3 that uses an adaptive tuning of its parameters, proposed by \cite{Seldin17}.
It is anytime, and it was also proven to obtain good ``best-of-both-world'' performance.
% meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.

>>>>

Similarly to what is considered for the first lower-bound,
a natural question is to know if there is an algorithm achieving a regret upper-bound of the form $R_T^{\cA}(I) \leq \bigO{\sqrt{K T}}$ for any instance, independently on the problem difficulty.
The question has been answered since the early $2000$s, as it is for instance the case for the \ExpThree{} algorithm from \cite{Auer02}.
In the last decade, some algorithms were shown to achieve both a problem-dependent logarithmic and a minimax upper-bounds,
like MOSS from \cite{Audibert2009minimax} or recently \KLUCBpp{} from \cite{Menard17},
and such results are usually referred to as ``best of both worlds''.

@@

% ----------------------------------------------------------------------------
\textbf{Policies for adversarial bandits.}
%
Another important family of bandits algorithms are the algorithms proposed for the adversarial setting \cite{Auer02NonStochastic}.
The first of them is the \ExpThree{} policy, for ``\emph{Exp}onential weights for \emph{Exp}loitation and \emph{Exp}loration'',
which is an example of the well-known \emph{multiplicative weights update algorithm}\footnote{~For more details, see
% on Wikipedia \href{https://en.wikipedia.org/wiki/Multiplicative_weight_update_method}{\texttt{en.wikipedia.org/wiki/Multiplicative\_weight\_update\_method}},
this article by J. Kun, \href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{JeremyKun.com/2017/02/27/the-reasonable-effectiveness-}}\\
\href{https://jeremykun.com/2017/02/27/the-reasonable-effectiveness-of-the-multiplicative-weights-update-algorithm/}{\texttt{of-the-multiplicative-weights-update-algorithm}}. The survey \cite{Arora2012multiplicative} gives more details.},
% which comes with an interactive online demonstration to help understanding the behavior of the algorithm, hosted at \href{https://j2kun.github.io/mwua/index.html}{\texttt{j2kun.GitHub.io/mwua/}} (accessed the 24th of May 2019)
a family of algorithms developed since the 1950s, and applied to a wide range of different domains.
% https://smpybandits.github.io/docs/Policies.Exp3.html
%
The \ExpThree{} policy maintains weights $w_k(t)$ on the $K$ arms, and at time $t$ it samples an arm from the distribution which is a mixture of $\bm{w}(t)=[w_1(t),\dots,w_K(t)]$ and the uniform distribution, \ie,
$\Pr(A(t)=k) \eqdef (1-\gamma) w_k(t) + \gamma \frac{1}{K}$.
The weights are initially uniform, \ie, $w_k(0)=1/K$, and then at each time after observing a reward $Y_{k,t}$ from arm $k$, \ExpThree{} updates the weight of the chosen arm multiplicatively, using $w_k(t+1) = w_k(t) \times \exp(\widehat{r_k}(t) / (\gamma N_k(t)))$. The weights are then renormalized, to have a sum $\sum_k w_k(t+1) = 1$.
This notation $\widehat{r_k}(t)$ means an unbiased estimator of the reward $r(t)=Y_{k,t}$, that is, $\widehat{r_k}(t) = r(t) / \Pr(A(t)=k)$.
%
Like for the $\varepsilon$-greedy policy, the parameter $\gamma$ can be constant, or can be a non-increasing sequence $(\gamma_t)_{t\in\N^*}$.
%
It was proven in \cite{Auer02NonStochastic} that \ExpThree{} with constant parameter $\gamma$ achieves $R_T = \bigO{\sqrt{K T \log(T)}}$ problem-independent regret,
while using a decreasing sequence, such as $\gamma_t = \sqrt{\log(K) / K} \log(t) / t$ gives an order-optimal problem-independent regret upper-bound of $\bigO{\sqrt{K T}}$.
A short proof of this results and more details on \ExpThree{} are given in \cite{Bubeck12}.
%
Many other policies have been proposed for adversarial bandits, like Follow-the-Perturbed-Leader or Follow-the-Regularized-Leader for linear adversarial bandits,
% but we do not study them in this thesis,
and we refer to Parts~III and VI of \cite{LattimoreBanditAlgorithmsBook} for more details.

@@

% https://smpybandits.github.io/docs/Policies.Exp3PlusPlus.html
Finally, it is worth mentioning the \Exppp{} algorithm, as a recent variant of \ExpThree{} that uses an adaptive tuning of its parameters, proposed by \cite{Seldin17}.
It is anytime, and it was also proven to obtain good ``best-of-both-world'' performance.
% meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.

### 2019-07-04 2-Chapters/3-Chapter/chapter3.tex
First few pages of Ch7
Using uniform notations for Exp3, Exp3.S, Exp3.R

A small script \texttt{configuration.py}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/configuration.html}{SMPyBandits.GitHub.io/docs/configuration.html}})
is used to import the arm classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Arms.html}{Arms}} module), the policy classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}} module) and define the problems and the experiments.
Choosing the algorithms to include is done by changing
the \texttt{configuration["policies"]} list in the \texttt{configuration.py} file.
For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{klUCB}} in \texttt{Policies} module) against the non-anytime variant $\klUCB^{++}$ algorithm (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{UCB}}) and Thompson sampling (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Thompson}}) with a Beta posterior (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Posterior.Beta}})).

@@

    \item
    %  Algorithm 3 is
    $\mathrm{Exp}3^{++}$ from \cite{Seldin17} (\texttt{Exp3PlusPlus}), using $\alpha=3$ and $\beta=256$ as advised.
    % It is a recent anytime variant of the $\mathrm{Exp3}$ algorithm, which was proven to obtain good ``best of both world'' performances.
    % meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.

@@

% FIXME
\newpage

@@

% FIXME
% \newpage

@@

@@

\newpage  % WARNING
\newpage  % WARNING
% ----------------------------------------------------------------------------
\section{Conclusion}
\label{sec:3:conclusion}

>>>>

A small script \texttt{configuration.py}
% (\texttt{\href{https://SMPyBandits.GitHub.io/docs/configuration.html}{SMPyBandits.GitHub.io/docs/configuration.html}})
is used to import the arm classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Arms.html}{Arms}} module), the policy classes (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.html}{Policies}} module) and define the problems and the experiments.
Choosing the algorithms to include is done by changing
the \texttt{configuration["policies"]} list in the \texttt{configuration.py} file.
For instance, one can compare the standard anytime \klUCB{} algorithm (class \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCB.html}{klUCB}} in \texttt{Policies} module) against the non-anytime variant \KLUCBpp{} algorithm (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.klUCBPlusPlus.html}{klUCBPlusPlus}}), and also \UCB{} with $\alpha=1$ (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.UCBalpha.html}{UCB}}) and Thompson sampling (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Thompson.html}{Thompson}}) with a Beta posterior (\texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.Posterior.Beta.html}{Posterior.Beta}})).

@@

    \item
    %  Algorithm 3 is
    \Exppp{} from \cite{Seldin17} (\texttt{Exp3PlusPlus}), using $\alpha=3$ and $\beta=256$ as advised.
    % It is a recent anytime variant of the \ExpThree{} algorithm, which was proven to obtain good ``best of both world'' performances.
    % meaning that it achieves $\cO(K\log(T)/\Delta^2)$ regret for problem-dependent bounds, and $\cO(\sqrt{KT})$ for problem-independent bounds.

@@

% % FIXME
% \newpage

@@

@@

% FIXME
% \vspace*{30pt}
\clearpage
% \newpage

@@

% FIXME?
% \newpage  % WARNING
% \newpage  % WARNING
% ----------------------------------------------------------------------------
\section{Conclusion}
\label{sec:3:conclusion}

### 2019-07-04 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Using uniform notations for Exp3, Exp3.S, Exp3.R

From a bandit perspective, it looks like each player is using a stochastic bandit algorithm (\UCB{} or \klUCB) when interacting with $K$ arms that give a feedback (the reward, and not the sensing information) that is far from being \iid{} from some distribution, due to the collisions.
%
As such, the algorithm does not appear to be well justified, and one may rather want to use adversarial bandit algorithms like $\mathrm{Exp3}$ \citep{Auer02NonStochastic}, that do not require a stochastic (\iid) assumption on arms.
%
However, we found out empirically that \Selfish{} is doing surprisingly well when using \UCB-like indexes, greatly outperforming \Selfish{} based on $\mathrm{Exp3}$,
like what we found in Chapter~\ref{chapter:4} in harder settings.

>>>>

From a bandit perspective, it looks like each player is using a stochastic bandit algorithm (\UCB{} or \klUCB) when interacting with $K$ arms that give a feedback (the reward, and not the sensing information) that is far from being \iid{} from some distribution, due to the collisions.
%
As such, the algorithm does not appear to be well justified, and one may rather want to use adversarial bandit algorithms like \ExpThree{} \citep{Auer02NonStochastic}, that do not require a stochastic (\iid) assumption on arms.
%
However, we found out empirically that \Selfish{} is doing surprisingly well when using \UCB-like indexes, greatly outperforming \Selfish{} based on \ExpThree,
like what we found in Chapter~\ref{chapter:4} in harder settings.

### 2019-07-04 2-Chapters/5-Chapter/chapter5.tex
Using uniform notations for Exp3, Exp3.S, Exp3.R

Their idea is to use the same skeleton as \RhoRand{} \cite{darak2016bayesian}, that is to assign a \emph{rank} to each player, and make players change their rank after any collision. But instead of selecting a new rank uniformly at random among $[M]$, they proposed to use a ``second-stage'' learning policy, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.
This is a very natural idea: use a bandit algorithm to balance the exploration/exploitation aspect of rank selection, instead of a random hoping.
This algorithm is named \rhoLearn, and it is implemented in SMPyBandits,
% \footnote{See \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}{\texttt{SMPyBandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}} for its documentation.},
where it can use Bayes-UCB, or any of the $65$ or more bandit algorithms available in the library.
%
Even if no theoretical guarantee backs up this idea of second-stage learning with Bayes-UCB, empirical simulations found that it can be efficient.
We illustrate this in Table~\ref{table:5:comparisonRhoRandRhoLearn} below, where we consider the same problem as described above in Section~\ref{sub:5:withoutSensing} (Table~\ref{table:5:meanRegretSimulationsNoSensing}).
We include different \rhoLearn{} algorithms, that uses \klUCB{}, Bayes-UCB or Exp3 for the second-stage learning,
and we include both the centralized multiple-play version of \klUCB{} and \MCTopM-\klUCB{} for comparison.

% DONE run the simulations with N=100 and not N=4
% for M in 3 6 9; do DEBUG=False SAVEALL=False NOPLOTS=True M=$M K=9 N_JOBS=-1 N=100 T=10000 make moremultiplayers; read; done
\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|ccc}
    \textbf{Algorithm} $\;$ \textbackslash $\;$ Number of players & $M=3$ & $M=6$ & $M=9$ \\
        \hline
        Centralized multiple-play \klUCB{} & $\mathbf{92 \pm 20}$ & $\mathbf{70 \pm 16}$ & $\mathbf{0}$ \\
        \hline
        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6639 \pm 1035$ \\
        \hline
        \rhoLearn-\klUCB{} + \klUCB{} & $546 \pm 190$ & $1172 \pm 295$ & $1416 \pm 347$ \\
        \rhoLearn-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1363 \pm 331$ \\
        \rhoLearn-\klUCB{} + Exp3 & $529 \pm 175$ & $1659 \pm 331$ & $5134 \pm 976$ \\
        \hline
        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 \pm 13$ \\
        % \hline
    \end{tabular}
    \caption{Comparing \RhoRand{} and \rhoLearn{} on a simple MP-MAB problem with $K=9$ arms.}
    \label{table:5:comparisonRhoRandRhoLearn}
% \end{footnotesize}
\end{table}

@@

Their proposal is using the Exp3 algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
As soon as the ``coordinator'' can effectively use one algorithm to decide the arms played by all the $M$ players, they show that the regret will grow as $R_T = \bigO{M^{4/3} K^{2/3} (\log(K))^{1/3} T^{2/3}}$ as shown in Theorem~4.1
(they used $K$ for $M$ the number of players, and $N$ for $K$ the number of arms).
This bound is much worse than the one obtained for the first article \cite{WeiSrivastava18Distributed}, but is more general, and it is much worse than the bound of $\bigO{\sqrt{K T \log(K)}}$ obtained for Exp3 for the single-player adversarial setting in \cite{Auer02NonStochastic}.
Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.
However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB{} (while scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}).

>>>>

Their idea is to use the same skeleton as \RhoRand{} \cite{darak2016bayesian}, that is to assign a \emph{rank} to each player, and make players change their rank after any collision. But instead of selecting a new rank uniformly at random among $[M]$, they proposed to use a ``second-stage'' learning policy, based on Bayes-UCB, to (try to) learn the best rank while still exploring each rank from time to time.
This is a very natural idea: use a bandit algorithm to balance the exploration/exploitation aspect of rank selection, instead of a random hoping.
This algorithm is named \rhoLearn, and it is implemented in SMPyBandits,
% \footnote{See \href{https://smpybandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}{\texttt{SMPyBandits.github.io/docs/PoliciesMultiPlayers.rhoLearn.html}} for its documentation.},
where it can use Bayes-UCB, or any of the $65$ or more bandit algorithms available in the library.
%
Even if no theoretical guarantee backs up this idea of second-stage learning with Bayes-UCB, empirical simulations found that it can be efficient.
We illustrate this in Table~\ref{table:5:comparisonRhoRandRhoLearn} below, where we consider the same problem as described above in Section~\ref{sub:5:withoutSensing} (Table~\ref{table:5:meanRegretSimulationsNoSensing}).
We include different \rhoLearn{} algorithms, that uses \klUCB{}, Bayes-UCB or \ExpThree{} for the second-stage learning,
and we include both the centralized multiple-play version of \klUCB{} and \MCTopM-\klUCB{} for comparison.

% DONE run the simulations with N=100 and not N=4
% for M in 3 6 9; do DEBUG=False SAVEALL=False NOPLOTS=True M=$M K=9 N_JOBS=-1 N=100 T=10000 make moremultiplayers; read; done
\begin{table}[ht]
% \begin{footnotesize}
    \centering
    \begin{tabular}{c|ccc}
    \textbf{Algorithm} $\;$ \textbackslash $\;$ Number of players & $M=3$ & $M=6$ & $M=9$ \\
        \hline
        Centralized multiple-play \klUCB{} & $\mathbf{92 \pm 20}$ & $\mathbf{70 \pm 16}$ & $\mathbf{0}$ \\
        \hline
        \RhoRand-\klUCB{} & $417 \pm 103$ & $2481 \pm 449$ & $6639 \pm 1035$ \\
        \hline
        \rhoLearn-\klUCB{} + \klUCB{} & $546 \pm 190$ & $1172 \pm 295$ & $1416 \pm 347$ \\
        \rhoLearn-\klUCB{} + Bayes-UCB & $561 \pm 219$ & $1204 \pm 394$ & $1363 \pm 331$ \\
        \rhoLearn-\klUCB{} + \ExpThree{} & $529 \pm 175$ & $1659 \pm 331$ & $5134 \pm 976$ \\
        \hline
        \MCTopM-\klUCB{} & $244 \pm 39$ & $401 \pm 55$ & $44 \pm 13$ \\
        % \hline
    \end{tabular}
    \caption{Comparing \RhoRand{} and \rhoLearn{} on a simple MP-MAB problem with $K=9$ arms.}
    \label{table:5:comparisonRhoRandRhoLearn}
% \end{footnotesize}
\end{table}

@@

Their proposal is using the \ExpThree{} algorithm from \cite{Auer02NonStochastic} on a combinatorial problem: they consider ``meta'' arms that are affectations of the $M$ players to $M$ distinct arms among the $K$ arms.
As soon as the ``coordinator'' can effectively use one algorithm to decide the arms played by all the $M$ players, they show that the regret will grow as $R_T = \bigO{M^{4/3} K^{2/3} (\log(K))^{1/3} T^{2/3}}$ as shown in Theorem~4.1
(they used $K$ for $M$ the number of players, and $N$ for $K$ the number of arms).
This bound is much worse than the one obtained for the first article \cite{WeiSrivastava18Distributed}, but is more general, and it is much worse than the bound of $\bigO{\sqrt{K T \log(K)}}$ obtained for \ExpThree{} for the single-player adversarial setting in \cite{Auer02NonStochastic}.
Note that this bound in $T^{2/3}$ is of the same order as the one given in \cite{Avner15}, for the \MEGA{} algorithm.
However, we note that the dependency in $M$ is surprisingly better for their result than for the regret upper-bound for \MCTopM-\klUCB{} (while scales in $M^3$ in Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB}).
